Crunching Large Graphs with Commodity Processors

### 2011-04-20 doc/hotpar2011/softxmt-hotpar2011.tex
wording change in QPI section
minor edits to intro

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. Tools to analyze social networks and query
semantic graphs with billions of entities exist
today~\cite{joslyn:hpsf, ediger:msda, brocheler:cosi, pegasus,
  pregel}, but graphs of interest to defense applications are expected
to have trillions of entities~\cite{hartley:mssg, kolda:imkd}. Speeding
up these applications at a low cost would have a large impact in how
we deal with large data-sets and make them even more valuable.

@@

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. They do not fit in a single commodity machine's
memory. They are difficult lay out with locality, since every vertex
needs to be near every other vertex. They are difficult to partition
in a balanced way~\cite{lang, leskovec}, leading to hotspots and load
imbalance. High-latency inter-node communication is a limiting factor
in scalability~\cite{yoo:parallelbfs}.

>>>>

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. While tools to analyze social networks and
query semantic graphs with billions of entities exist
today~\cite{joslyn:hpsf, ediger:msda, brocheler:cosi, pegasus,
  pregel}, graphs of interest to defense applications are expected to
have trillions of entities~\cite{hartley:mssg, kolda:imkd}. Speeding
up these applications at a low cost would have a large impact in how
we deal with large data-sets and make them even more valuable.

@@

The most interesting computational challenge comes from large,
low-diameter, power law graphs: this combination makes extracting
performance difficult. They do not fit in a single commodity machine's
memory. They are difficult lay out with locality, since every vertex
needs to be near every other vertex. They are difficult to partition
in a balanced way~\cite{lang, leskovec}, leading to hotspots and load
imbalance. Consequently, high-latency inter-node communication becomes
an important limiting factor in scalability~\cite{yoo:parallelbfs}.
